\textcolor{blue}{A gradient-based optimization method for motion planning}$^{1}$

### 2016-01-27 paper.tex
Finish results part, fix Figures

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}} % \x already defined in the class
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
%\newcommand\tcolli{t_{coll\ i}}
\newcommand\tcolli{\kappa_i}
\newcommand\po{\mathbf{P}}
\newcommand\Jf{\Phi}
\newcommand\kernel{\mbox{Ker }}
\newcommand\U{\mathbf{u}}
\newcommand\rotation{R}
\newcommand\traj{\Gamma}
\newcommand\velocity{\mathbf{v}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\graphicspath{{images/}}

@@

For this shortening purpose, random shortcut (RS) methods are
still very 
popular~\cite{Sekhavat-Svestka1998,GeraertsIJRR07,HauserFastSmooth}. 
However, RS requires fine 
tuning of the termination condition and is no efficient for long 
trajectories where only a minor part needs to be optimized, 
as in \textcolor{blue}{Figure~\ref{local_box_optim}}.
Figure~\ref{decoupled_DOF_optimization} presents another situation where RS 
will always fail to optimize the initial path, since it cannot decouple the 
robot degrees of freedom (DOF) on which the optimization occurs. This problem is 
addressed by our method.

@@

The idea of our method is to find a good trade-off between
the simplicity of “blind” methods like shortcut algorithms,
and the complexity of distance based optimization techniques.
The method iteratively shortens the initial path with gradient-based information.
When a collision is detected at a given iteration, the method backtracks to the
latest valid iteration and inserts a one-dimensional constraint
between the objects detected in collision. Only collisions between objects are 
evaluated, therefore no pre-processing of either the
robot or environment models is necessary to increase distance computation speed. 
Respecting the problem geometry also 
preserves that a solution can still be found, e.g. for narrow passages as holes or 
grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a \textcolor{blue}{linearly constrained} quadratic program.

\textcolor{blue}{Another important feature of our contribution is that we optimize paths on the robot configuration space in a proper mathematical way. Most other optimization algorithms represent $SO3$ rotations by a vector directed along the rotation axis and the norm of which is the rotation angle, also known as the exponential map of $so3$, or even worse by Euler angles.}

@@

\textcolor{blue}{This section describes the settling of the Gradient-based optimizer. The method works as a classical Linearly Constrained Quadratic Program (LCQP), reducing the path length with a cost function and avoid collisions with constraints. Details of the LCQP elements will be given in the following subsections, and associated to functions that will populate the global algorithm, presented in the last section.}

@@

Each path $\xx$ is a mapping from 
interval $[0,1]$ into $\CS$: $\xx(0) = q_0$, $\xx(1) = q_{wp+1}$.
Finally, we use a continuous 
collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
paths and to return the first colliding configuration 
\textcolor{blue}{and its abcissa} along a path. \textcolor{blue}{This stage 
is handled by a function \texttt{validatePath}, which returns $true$ if the given path is collision-free}.

@@

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{red}{We denote by
  \begin{itemize}
  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
$$
\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - P_1^{loc}\|}
$$
Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
}
\textcolor{blue}{
  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
\begin{equation}\label{eq:g}
g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - P_1^{loc} | \U\right)
\end{equation}
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
\begin{equation}\label{eq:f}
f (\xx) = g(\xx (\tcolli))
\end{equation}
\textcolor{red} {The constraint defined for any path $\xx$ by:
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
aims at keeping point $\po_2$ in a plane fixed to $\body_1$, orthogonal to $\U$ and and passing by $\po_2$ in configuration $\xx_i(\tcolli)$ (Figure~\ref{contact-points} right).
}
We linearize the constraint around $\xx_{i}$: 
$$
\frac{\partial f}{\partial \xx}(\xx_i)(\xx - \xx_i) = 0
$$
The computation of the linearized constraint is described in the next
section. Then, a line is added in the constraint Jacobian matrix $\Jf_i$:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{j+1}\end{array}\right)\ \mbox {with}\\
L_{j+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}
\textcolor{blue}{This step is performed by function \texttt{addCollisionConstraint}}.

@@

\begin {figure}
  \centerline {
    \def\svgwidth {.7\linewidth}
    \graphicspath{{./images/}}
    \input {images/convergence.pdf_tex}
  }
  \caption {\textcolor{red}{Illustration of linearly constrained quadratic program on a problem analogous to the path shortening problem: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2, \xx\in\real^2$ under the non-linear constraint $f(\xx)\leq 0$. $\alpha_{init}=\frac{1}{4}$. The algorithm starts from $\xx_0$. The first iterate is $\xx_1$ which satisfies the constraint. The second iterate is $\xx_2$ that does not satisfy the constraint. The algorithm backtracks to $\xx_1$ and inserts linear constraint $L_1: \frac{\partial f}{\partial \xx} (\xx_1)(\xx - \xx_1) = 0$. $\xx_3$ is the global minimum under $L_1$. As $\xx_3$ does not satisfy the constraint, the algorithm moves to $\xx_4$ that satisfies the constraints, and then to $\xx_5$ that does not. The algorithms backtracks to $\xx_4$, inserts constraint $L_2: \frac{\partial f}{\partial \xx} (\xx_2)(\xx - \xx_2) = 0$, and returns $\xx_4$ as a solution since the dimension of the search space is 0. Notice that the same non-linear constraint may give rise to several linear constraints. The convergence of the algorithm relies on the fact that the kernel of constraint $L2$ is not contained in the kernel of the current constraints ($L_1$ only here). The convergence analysis can be roughly summarized as follows. $L2$ to be linearly dependent of $L_1$ requires that $f$ is stationary along $L_1$ at $\xx_4$. This is unlikely (but possible) since $\xx_4$ is not far from the boundary of the domain defined by $f(\xx) \leq 0$. If $L_2$ was not linearly independent from $L_1$, The algorithm would keep searching new iterates between $\xx_5$ and $\xx_4$. If by any chance constraint $f$ linearized around each of those collision-free iterates was each time linearly dependent from $L_1$ the iterates would converge to the boundary of the domain defined by $f(\xx) \leq 0$. By continuous differentiability of $f$, this would mean that $f$ is stationary at the boundary. In other words, the straight line passing by $\xx_1$ and $\xx_3$ would cross the boundary tangently (as in bottom right picture). This is possible but unlikely, unless the problem has been defined as such on purpose.}}
  \label {fig:convergence}
\end{figure}

@@

\subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
\noindent
\textcolor{red}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
According to~(\ref{eq:f}), the evaluation of the constraint function $f$ along the iteration line $\xx_i + t\alpha_i \p_i,\ t\in [0,1]$ going from $\xx_i$ to $\xx_{i+1}$ can be written as follows:
\begin{equation}\label{eq:f-along-pi}
f (\xx_i + t\alpha_i \p_i) = g \left((\xx_i + t\alpha_i \p_i)(\tcolli)\right).
\end{equation}
The argument of function $g$ above is a trajectory in the robot configuration space that we denote by $\traj$:
\begin{equation}\label{eq:trajectory}
\traj (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
\end{equation}
The trajectory $\traj(t)$ is defined by taking the constant abscissa $\tcolli$ and by 
moving from 
path $\xx_i$ along step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
Substituting~(\ref{eq:trajectory}) into~(\ref{eq:f-along-pi}) and differentiating with respect  to $t$ yields
\begin{align}
f (\xx_i + t\alpha_i \p_i) &= g (\traj (t)) \\
\label{eq:df/dx}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i + t\alpha_i \p_i)\p_i &=
\frac{d}{dt}g (\traj (t))
\end{align}
From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $P_2$ in reference frame $\body_1$ projected on vector $\U$.
}

\textcolor{red}{
Therefore the following expressions
\begin{align*}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
\end{align*}
correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $P_2$ in reference frame $\body_1$. 
Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0. According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$. Although very unlikely, this case might appear for some new constraint. In this case, inserting constraint $L_{j+1}$ is useless since
$$
\kernel \Jf_i = \kernel \Jf_{i+1}.
$$
}

@@

\textcolor{red}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent. This loop is performed by Algorithm~\ref{algo:solveRedundant} and illustrated in Figure~\ref{fig:convergence-diagram}. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
}

\begin{figure}
	\centering
	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
	\caption{\textcolor{blue}{Considering a trajectory 
	$\conf (t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
	and $\kappa$ respectively collision-free path, in-collision path, iteration step 
	and collision-abcissa).
	The trajectory is represented in the configuration space (left) and in the 
	workspace with the colliding bodies (right).}
	Note that a case with 
	$(\velocity_{\po_2/\body_1}(\xx_{Free})|\mathbf{u})=0$ is considered, so $\p$ already complies 
	with the constraint that should be added if no redundancy was tracked. Therefore, 
	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
	$(\velocity_{\po_2/\body_1}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
	returning the new constraint.}
	\label{fig:convergence-diagram}
\end{figure}

\begin{algorithm}[H]
\begin{algorithmic}%[1] % line numerotation
\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
\OUTPUT{linearized constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$}
\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
\State $solved \gets false$
\While{($\textbf{not}(solved)$)}
	%\State $\p = \texttt{computeIterate}()$
	%\State $minReached = (||\p||<10^{-3} \textbf{ or } \alpha=1)$
	\State $\xx \gets \xx_{Free} + 0.5\alpha\,\p$
	\If{(\texttt{validatePath}($\xx$))}
		\State $\xx_{Free} \gets \xx$
	\Else
		\State $\xx_{Coll} \gets \xx$
	\EndIf
	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
	\textcolor{blue}{\State $\alpha \gets 0.5\alpha$}
\EndWhile
\end{algorithmic}
\caption{Description of \texttt{findNewConstraint}() which returns a linearly 
independent constraint w.r.t. previous constraints stacked in $\Jf$.} 
\label{algo:solveRedundant}
\end{algorithm}

@@

\vspace{0.4cm}
% freeflyer puzzle
Before entering the manipulator examples, the GB algorithm is analyzed 
on a popular problem in the motion planning literature: a freeflyer puzzle, 
corresponding to Figure~\ref{fig:trajectories:puzzle}. The 
puzzle has to cross down the 
obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
contains detours above and below the obstacle, as well as small superfluous
motions in the hole. 
The GB optimizer manages to modify the puzzle trajectory even in the hole, but remains less efficient than RS and PRS. We explain this behavior by the fact that, ........
%PRS fails to decrease the path length because in this problem, all DOF are closely related to cross the hole. This behavior was also noticed and improved in the PRS implementation of~\cite{GeraertsIJRR07}, acting on a group of DOF. The GB optimizer manages to modify the puzzle trajectory even in the hole, contrary to RS which simply copies the initial path when the trajectory gets near of the obstacle.

>>>>

% Math shortcuts :
\newcommand\real{\mathbb{R}}
\newcommand\p{\mathbf{p}}
\newcommand\gT{\tilde{g}^T}
\newcommand\g{\tilde{g}}
\newcommand\CS{\mathcal{C}}
\newcommand\dimCS{n_\mathbf{C}}
\newcommand\body{{\cal B}}
\newcommand\Sone{\mathbf{S}^1}
\newcommand\Sthree{\mathbf{S}^3}
\newcommand\conf{\mathbf{q}}
\newcommand\xx{\mathbf{x}} % \x already defined in the class
\newcommand\cost{C}
\newcommand\weight{W}
\newcommand\translation{\mathbf{t}}
%\newcommand\tcolli{t_{coll\ i}}
\newcommand\tcolli{\kappa_i}
\newcommand\po{\mathbf{P}}
\newcommand\Jf{\Phi}
\newcommand\kernel{\mbox{Ker }}
\newcommand\U{\mathbf{u}}
\newcommand\rotation{R}
\newcommand\traj{\Gamma}
\newcommand\velocity{\mathbf{v}}
\newcommand\todo{\textcolor{red}{NOT FINISHED}}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\graphicspath{{images/}}

@@

For this shortening purpose, random shortcut (RS) methods are
still very 
popular~\cite{Sekhavat-Svestka1998,GeraertsIJRR07,HauserFastSmooth}. 
However, RS requires fine 
tuning of the termination condition and is no efficient for long 
trajectories where only a minor part needs to be optimized, 
as in \textcolor{blue}{Figure~\ref{local_box_optim}}.
Figure~\ref{decoupled_DOF_optimization} presents another situation where RS 
will always fail to optimize the initial path, since it cannot decouple the 
robot degrees of freedom (DOF) on which the optimization occurs. This problem is 
addressed by our method.
\textcolor{blue}{Processing a path 
pruning~\cite{GeraertsIJRR07}, in order to remove redundant nodes from the initial 
path, is a classical preliminary step for path length shortening. 
A pruning will efficiently solve the Figure~\ref{local_box_optim} issue, however it 
will fail tackling the Figure~\ref{decoupled_DOF_optimization} one, as RS.}

@@

The idea of our method is to find a good trade-off between
the simplicity of “blind” methods like shortcut algorithms,
and the complexity of distance based optimization techniques.
The method iteratively shortens the initial path with gradient-based information.
When a collision is detected at a given iteration, the method backtracks to the
latest valid iteration and inserts a one-dimensional constraint
between the objects detected in collision. Only collisions between objects are 
evaluated, therefore no pre-processing of either the
robot or environment models is necessary to increase distance computation speed. 
Respecting the problem geometry also 
preserves that a solution can still be found, e.g. for narrow passages as holes or 
grippers. The method is also repeatable since no randomness is introduced. The underlying optimization algorithm is a \textcolor{blue}{Linearly Constrained Quadratic Program (LCQP)}.

\textcolor{blue}{Another important feature of our contribution is that we optimize paths on the robot configuration space in a proper mathematical way. Most other optimization algorithms represent $SO(3)$ rotations by a vector directed along the rotation axis and the norm of which is the rotation angle, also known as the exponential map of $SO(3)$, or even worse by Euler angles.}

@@

\textcolor{blue}{This section describes the establishment of the Gradient-based optimizer. The method works as a classical Linearly Constrained Quadratic Program (LCQP), reducing the path length expressed as a cost function and avoid collisions with linearized constraints. Details of the LCQP elements will be given in the following subsections, and associated to functions that will populate the global algorithm, presented in the last section.}

@@

Each path $\xx$ is a mapping from 
interval $[0,1]$ into $\CS$: $\xx(0) = q_0$, $\xx(1) = q_{wp+1}$.
Finally, we use a continuous 
collision checker inspired of~\cite{SchwarzerExactCollision} to validate our 
paths and to return the first colliding configuration 
\textcolor{blue}{and its abcissa} along a path.

@@

In the collision-configuration $\xx_{i+1}(\tcolli)$, let $\po_c\in \real^3$ be a 
contact point expressed in the global frame 
(Figure~\ref{contact-points} left).
\textcolor{blue}{We denote by
  \begin{itemize}
  \item $\po_1^{loc}$ (resp. $\po_2^{loc}$) the coordinate vector of $\po_c$ in the local frame of  $\body_1$ (resp. $\body_2$),
  \item $M_1(\conf) \in SE(3)$ (resp. $M_2(\conf) \in SE(3)$) the rigid-body transformation representing the position of $\body_1$ (resp. $\body_2$) in the global frame in configuration $\conf$,
  \item $M_2^1 (\conf) = M_1(\conf)^{-1} M_2(\conf)$ the position of $\body_2$ local frame in $\body_1$ local frame in configuration $\conf$,
  \item $\po_1(\conf)$ (resp. $\po_2(\conf)$) the points moving with $\body_1$ (resp. $\body_2$) of local coordinate $\po_1^{loc}$ (resp. $\po_2^{loc}$) in $\body_1$ (resp. $\body_2$) local frame.
  %% \item $\rotation_1(\conf) \in SO(3)$ (resp. $\rotation_2(\conf) \in SO(3)$) the rotation associated with $M_1(\conf)$ (resp. $M_2(\conf)$).
  \end{itemize}
We define $\U$ as the coordinate vector of the unit vector linking points $\po_1$ and $\po_2$ in configuration $\xx_{i}(\tcolli)$, expressed in local frame of $\body_1$ (Figure~\ref{contact-points} right):
$$
\U = \frac{M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}}{\|M_2^1 (\xx_{i}(\tcolli))\po_2^{loc} - \po_1^{loc}\|}
$$
Note that $\U$ is well defined since configuration $\xx_{i}(\tcolli)$ is collision-free.
}
\textcolor{blue}{
  Let $g$ be the real valued function mapping the projection of vector $\overrightarrow{\po_1\,\po_2}$ on $\mathbf{u}$ to a configuration $\conf$ . For any $\conf \in \CS$:
\begin{equation}\label{eq:g}
g (\conf) = \left(M_2^1 (\conf)\po_2^{loc} - \po_1^{loc} | \U\right)
\end{equation}
}
%which is calculated from the last collision-free path at abscissa $\tcolli$.

Let $f$ be the function defined from $\CS^{wp}$ to $\real$ by:
\begin{equation}\label{eq:f}
f (\xx) = g(\xx (\tcolli))
\end{equation}
\textcolor{blue} {The constraint defined for any path $\xx$ by:
\begin {equation}\label{eq:new-constraint}
f(\xx) - f(\xx_{i}) = 0.
\end {equation}
aims at keeping point $\po_2$ in a plane fixed to $\body_1$, orthogonal to $\U$ and and passing by $\po_2$ in configuration $\xx_i(\tcolli)$ (Figure~\ref{contact-points} right).
}
We linearize the constraint around $\xx_{i}$: 
$$
\frac{\partial f}{\partial \xx}(\xx_i)(\xx - \xx_i) = 0
$$
The computation of the linearized constraint is described in the next
section. Then, a line is added in the constraint Jacobian matrix $\Jf_i$:
\begin {align*}
\Jf_{i+1} &= \left(\begin{array}{c}L_1 \\ \vdots \\ L_{j+1}\end{array}\right)\ \mbox {with}\\
L_{j+1} &= \frac{\partial f}{\partial \xx}(\xx_i)
\end{align*}
\textcolor{blue}{This stage is performed by a function \texttt{addCollisionConstraint}}.

@@

\begin {figure}
  \centerline {
    \def\svgwidth {.7\linewidth}
    \graphicspath{{./images/}}
    \input {images/convergence.pdf_tex}
  }
  \caption {\textcolor{blue}{Illustration of linearly constrained quadratic program on a problem analogous to the path shortening problem: we want to minimize the quadratic cost $\frac{1}{2}\|\xx-\xx^*\|^2, \xx\in\real^2$ under the non-linear constraint $f(\xx)\leq 0$. $\alpha_{init}=\frac{1}{4}$. The algorithm starts from $\xx_0$. The first iterate is $\xx_1$ which satisfies the constraint. The second iterate is $\xx_2$ that does not satisfy the constraint. The algorithm backtracks to $\xx_1$ and inserts linear constraint $L_1: \frac{\partial f}{\partial \xx} (\xx_1)(\xx - \xx_1) = 0$. $\xx_3$ is the global minimum under $L_1$. As $\xx_3$ does not satisfy the constraint, the algorithm moves to $\xx_4$ that satisfies the constraints, and then to $\xx_5$ that does not. The algorithms backtracks to $\xx_4$, inserts constraint $L_2: \frac{\partial f}{\partial \xx} (\xx_2)(\xx - \xx_2) = 0$, and returns $\xx_4$ as a solution since the dimension of the search space is 0. Notice that the same non-linear constraint may give rise to several linear constraints. The convergence of the algorithm relies on the fact that the kernel of constraint $L2$ is not contained in the kernel of the current constraints ($L_1$ only here). The convergence analysis can be roughly summarized as follows. $L2$ to be linearly dependent of $L_1$ requires that $f$ is stationary along $L_1$ at $\xx_4$. This is unlikely (but possible) since $\xx_4$ is not far from the boundary of the domain defined by $f(\xx) \leq 0$. If $L_2$ was not linearly independent from $L_1$, The algorithm would keep searching new iterates between $\xx_5$ and $\xx_4$. If by any chance constraint $f$ linearized around each of those collision-free iterates was each time linearly dependent from $L_1$ the iterates would converge to the boundary of the domain defined by $f(\xx) \leq 0$. By continuous differentiability of $f$, this would mean that $f$ is stationary at the boundary. In other words, the straight line passing by $\xx_1$ and $\xx_3$ would cross the boundary tangently (as in bottom right picture). This is possible but unlikely, unless the problem has been defined as such on purpose.}}
  \label {fig:convergence}
\end{figure}

@@

\subsubsection{Note about assumption~(\ref{eq:assumption-1})}%
\noindent
\textcolor{blue}{As in the previous section, we assume that $\xx_i$ is collision-free and that $\xx_{i+1}$ is in collision at abscissa $\tcolli$.
According to~(\ref{eq:f}), the evaluation of the constraint function $f$ along the iteration line $\xx_i + t\alpha_i \p_i,\ t\in [0,1]$ going from $\xx_i$ to $\xx_{i+1}$ can be written as follows:
\begin{equation}\label{eq:f-along-pi}
f (\xx_i + t\alpha_i \p_i) = g \left((\xx_i + t\alpha_i \p_i)(\tcolli)\right).
\end{equation}
The argument of function $g$ above is a trajectory in the robot configuration space that we denote by $\traj$:
\begin{equation}\label{eq:trajectory}
\traj (t) = (\xx_{i} + t\, \alpha_i \p_{i}) (\tcolli), \ t \in [0,1]
\end{equation}
The trajectory $\traj(t)$ is defined by taking the constant abscissa $\tcolli$ and by 
moving from 
path $\xx_i$ along step $\p_i$. Note that configuration $\xx_{i+1}(\tcolli)$ is reached when $t=1$.
Substituting~(\ref{eq:trajectory}) into~(\ref{eq:f-along-pi}) and differentiating with respect  to $t$ yields
\begin{align}
f (\xx_i + t\alpha_i \p_i) &= g (\traj (t)) \\
\label{eq:df/dx}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i + t\alpha_i \p_i)\p_i &=
\frac{d}{dt}g (\traj (t))
\end{align}
From the definition~(\ref{eq:g}) of $g$, the right hand side of~(\ref{eq:df/dx}) represents the velocity of point $\po_2$ in reference frame $\body_1$ projected on vector $\U$.
}

\begin{figure}
	\centering
	\includegraphics[width=15.8cm]{convergence-diagram.pdf}
	\caption{\textcolor{blue}{Considering a trajectory 
	$\traj(t) = (\xx_{Free} + t\, \alpha \p)(\kappa)$ 
	with $\xx_{Coll} = \xx_{Free} + \alpha\p$ ($\xx_{Free}$, $\xx_{Coll}$, $\p$ 
	and $\kappa$ are respectively collision-free path, in-collision path, 
	iteration step and collision-abcissa).
	The trajectory is represented in the configuration space (left) and in the 
	workspace with the colliding bodies (right).}
	Note that a case with 
	$(\velocity_{\po_2/\body_1}(\xx_{Free})|\mathbf{u})=0$ is considered, 
	so $\p$ already complies 
	with the constraint that should be added if no redundancy was 
	tracked. Therefore, 
	Algorithm~\ref{algo:solveRedundant} will produce one intermediate step
	$\xx$ (gray path and clear bodies). $\xx$ is collision-free and 
	$(\velocity_{\po_2/\body_1}(\xx)|\mathbf{u})\neq 0$ so the redundancy is solved, 
	returning the new constraint.}
	\label{fig:convergence-diagram}
\end{figure}

\textcolor{blue}{
Therefore the following expressions
\begin{align*}
\alpha_i \frac{\partial f}{\partial \xx} (\xx_i)\p_i = \frac{d}{dt}g (\traj (0)) & &
\alpha_i \frac{\partial f}{\partial \xx} (\xx_{i+1})\p_i =\frac{d}{dt}g (\traj (1))
\end{align*}
correspond to $(\velocity_{\po_2/\body_1}|\mathbf{u})$, respectively in configurations $\xx_{i}(\tcolli)$ and $\xx_{i+1}(\tcolli)$, where $\velocity_{\po_2/\body_1}$ represents the velocity of point $\po_2$ in reference frame $\body_1$. 
Note that Assumption~(\ref{eq:assumption-1}) is equivalent to the first above equality to be different from 0. According to the geometric interpretation, Assumption~(\ref{eq:assumption-1}) is violated if and only if $\velocity_{\po_2/\body_1}$ is orthogonal to $\U$. Although very unlikely, this case might appear for some new constraint. In this case, inserting constraint $L_{j+1}$ is useless since
$$
\kernel \Jf_i = \kernel \Jf_{i+1}.
$$
}

\begin{algorithm}
\begin{algorithmic}%[1] % line numerotation
\INPUT{$\xx_{Free}$ latest collision-free path, $\xx_{Coll}$ current path with collisions, $\p$ current iteration step}
\OUTPUT{linearized constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$}
\Require constraint $\frac{\partial f}{\partial \xx} (\xx_{Free})$ built from $\xx_{Coll}$ would produce a rank loss in constraint Jacobian $\Jf$
\State $solved \gets false$
\While{($\textbf{not}(solved)$)}
	%\State $\p = \texttt{computeIterate}()$
	\State $\alpha \gets 0.5\alpha$
	\State $\xx \gets \xx_{Free} + \alpha\,\p$
	\If{(\texttt{validatePath}($\xx$))}
		\State $\xx_{Free} \gets \xx$
	\Else
		\State $\xx_{Coll} \gets \xx$
	\EndIf
	\State $\frac{\partial f}{\partial \xx} (\xx_{Free}) \gets$\texttt{computeCollisionConstraint}($\xx_{Coll}$, $\xx_{Free}$)
	\State $solved \gets$ \texttt{isFullRank}($\Jf$ , $\frac{\partial f}{\partial \xx} (\xx_{Free})$)
\EndWhile
\end{algorithmic}
\caption{Description of \texttt{findNewConstraint}() which returns a linearly 
independent constraint w.r.t. previous constraints stacked in $\Jf$.} 
\label{algo:solveRedundant}
\end{algorithm}

@@

\textcolor{blue}{
When the new constraint is not linearly independent from the set of previous constraints, the algorithm enters an additional loop in order to find a new constraint that is linearly independent. This loop is performed by Algorithm~\ref{algo:solveRedundant} and illustrated in Figure~\ref{fig:convergence-diagram}. The loop keeps looking for paths along line segment $[\xx_{i},\xx_{i+1}]$ by dichotomy. A pair containing the latest free path and the latest path in collision, denoted by $(\xx_{Free}, \xx_{Coll})$ is stored along the loop. New iterations are chosen in the middle of this pair. 
\begin{itemize}
\item If the new path is collision-free, it replaces $\xx_{Free}$ in the pair.
\item If the new path is in collision, it replaces $\xx_{Coll}$ in the pair.
\end{itemize}
In both cases, a new constraint is built following the method described in Section~\ref{subsubsec:new-constraint}. Two cases are then possible:
\begin{enumerate}
\item at some point in the loop the new constraint is linearly independent from the previous constraints. The new constraint is added to $\Jf_{i}$ to give rise to $\Jf_{i+1}$ and the loop is interrupted, or
\item each new constraint is linearly dependent from the previous constraints and the loop never ends.
\end{enumerate}
In the second case, the iterations of the loop converge to a path that we denote by $\bar{\xx}$. $\xx_{Free}$ and $\xx_{Coll}$ also both converge to $\bar{\xx}$. Note that $\bar{\xx}$ lies at the boundary between free paths and paths in collision along the search line.
}

\textcolor{red}{Partie en cours de modif par Florent}

@@

\vspace{0.4cm}
% freeflyer puzzle
Before entering the manipulator examples, the GB algorithm is analyzed 
on a popular problem in the motion planning literature: a freeflyer puzzle, 
corresponding to Figure~\ref{fig:trajectories:puzzle}. The 
puzzle has to cross down the 
obstacle using the hole in the middle. The initial path planned with Visibility-PRM 
contains detours above and below the obstacle, as well as small superfluous
motions in the hole. 
The GB optimizer manages to modify the puzzle trajectory even in the hole, but remains less efficient than RS and PRS to shorten trajectories in the upper and lower parts. 
\textcolor{blue}{
This can be the result of adding collision-constraints on these parts of the trajectories, penalizing the global path length. One idea could be to 	arbitrarily cancel constraints in upper and lower parts, and keeping the ones in the hole. However we want the present GB algorithm to remain general and basic, such constraints relaxation are part of the future possible work.
}
%PRS fails to decrease the path length because in this problem, all DOF are closely related to cross the hole. This behavior was also noticed and improved in the PRS implementation of~\cite{GeraertsIJRR07}, acting on a group of DOF. The GB optimizer manages to modify the puzzle trajectory even in the hole, contrary to RS which simply copies the initial path when the trajectory gets near of the obstacle.
